---
title: "chickenpox"
author: "MaryJoWebster"
date: "March 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

#create tables with varicella (chicken pox) vaccination data only
#separate tables for data year (12-13 and 17-18) and grade (kindergarten and 7th)
var_kind_1718 <-  kind_new %>% select(schoolID, disname, schname, enroll_new, 
                                 complete_var, inprogress_var, co_var, me_var,
                                 complete_pert_var, inprogress_pert_var, co_pert_var, me_pert_var, disease_hist_var, diseasehist_pert_var,
                                 pctpoverty, pctminority, income_bucket, 
                                 diversity_bucket,  
                                 county, type2, yr, schoolyr, k12enr, grade, schooltype)

var_kind_1213 <-  kind_old  %>% select(schoolID, disname, schname, enroll_new, 
                                  complete_var, inprogress_var, co_var, me_var,
                                  complete_pert_var, inprogress_pert_var, co_pert_var, me_pert_var,  
                                  pctpoverty, pctminority, income_bucket, 
                                  diversity_bucket,  
                                  county, type2, yr, schoolyr, k12enr, grade, schooltype)

var_7_1718 <-  seventh_new %>% select(schoolID, disname, schname, enroll_new, 
                                      complete_var, inprogress_var, co_var, me_var,
                                      complete_pert_var, inprogress_pert_var, co_pert_var, me_pert_var, disease_hist_var, diseasehist_pert_var,
                                      pctpoverty, pctminority, income_bucket, 
                                      diversity_bucket,  
                                      county, type2, yr, schoolyr, k12enr, grade, schooltype)

var_7_1213 <-  seventh_old  %>% select(schoolID, disname, schname, enroll_new, 
                                       complete_var, inprogress_var, co_var, me_var,
                                       complete_pert_var, inprogress_pert_var, co_pert_var, me_pert_var,  
                                       pctpoverty, pctminority, income_bucket, 
                                       diversity_bucket,  
                                       county, type2, yr, schoolyr, k12enr, grade, schooltype)


#merge the files together
var <-  bind_rows(var_kind_1213, var_kind_1718, var_7_1718, var_7_1213)

#populate null values with zero
var$co_var[is.na(var$co_var)] <- 0
var$complete_var[is.na(var$complete_var)] <- 0
var$inprogress_var[is.na(var$inprogress_var)] <- 0
var$me_var[is.na(var$me_var)] <- 0
var$disease_hist_var[is.na(var$disease_hist_var)] <- 0
var$co_pert_var[is.na(var$co_pert_var)] <- 0
var$complete_pert_var[is.na(var$complete_pert_var)] <- 0
var$inprogress_pert_var[is.na(var$inprogress_pert_var)] <- 0
var$me_pert_var[is.na(var$me_pert_var)] <- 0
var$diseasehist_pert_var[is.na(var$diseasehist_pert_var)] <- 0



#calculate percent immunity
#this combines the number kids vaccinated, plus those who had the disease previously
var <-  var %>%     mutate( 
  total_immunity = complete_var+disease_hist_var,
  total_kids = total_immunity+inprogress_var+co_var+me_var,
  pct_immunity=total_immunity/total_kids,
    var_pocket=case_when(
  enroll_new==0~'not reported',
  pct_immunity>=.895~'no',
                                                            pct_immunity>=0 & pct_immunity<.895~'yes',
                                                            is.na(pct_immunity)~'not reported'))



#assign the primary "reason" schools have a low var rate
#looks at whether school has higher rate of incomplete or higher rate of waivers
var <-  var %>% mutate(pocket_reason = case_when(var_pocket=='no'~'na',
                                                 var_pocket=='yes' & co_pert_var>inprogress_pert_var~'non-medical waiver',
                                                 var_pocket=='yes' & inprogress_pert_var>co_pert_var~'in progress',
                                                 var_pocket=='yes' & co_pert_var==inprogress_pert_var~'equal',
                                                 var_pocket=='not reported'~'not reported',
                                                 TRUE~'not clear'))





```

#Pct of schools in 2018 lacking immunity
pocket=yes means school has less than 90% immune to chicken pox<br>
This combines seventh and kindergarten
```{r}
  var%>% filter(yr==2018, var_pocket!='not reported') %>% tabyl(grade, var_pocket)%>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  adorn_title("top")%>%
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```
#Pct of schools in 2018 lacking immunity-Kindergarten
By type of school<br>
pocket=yes means school has less than 90% immune to chicken pox<br>
```{r}
  var%>% filter(yr==2018, var_pocket!='not reported',  grade=='Kindergarten') %>%  tabyl(type2, var_pocket)%>%

  adorn_percentages("row") %>%
   adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  adorn_title("top")%>%
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

#Pocket schools by primary reason
Both kindergarten and 7th grade, 2018 only
```{r}
var %>% filter(yr==2018, var_pocket=='yes') %>% 
  group_by(pocket_reason) %>% 
  summarize(count=n()) %>% 
  mutate(pct=count/sum(count))
```

#2018, kindergarten summary by type of school
```{r}
var %>% filter(yr==2018, var_pocket!='not reported', grade=='Kindergarten') %>% 
  group_by(type2) %>% 
  summarise(complete = sum(total_immunity), 
            partial = sum(inprogress_var),
            nonmed = sum(co_var),
            med = sum(me_var),
            totkids = sum(total_kids)) %>% 
  mutate(pctcomplete=complete/totkids,
         pctwaiver = nonmed/totkids) %>% 
    knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```
#2013, kindergarten summary by grade
```{r}
var %>% filter(yr==2013, var_pocket!='not reported') %>% 
  group_by(grade) %>% 
  summarise(complete = sum(total_immunity), 
            partial = sum(inprogress_var),
            nonmed = sum(co_var),
            med = sum(me_var),
            totkids = sum(total_kids)) %>% 
  mutate(pctcomplete=complete/totkids,
         pctwaiver = nonmed/totkids) %>% 
    knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

