---
title: "vaccination_analysis"
author: "MaryJoWebster"
date: "March 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, code=readLines("process_vaccination_data.R"), echo=FALSE, warning=FALSE, message=FALSE}

```
#MMR analysis<br>

Analysis includes both kindergarten and 7th grade unless noted

<br>


#Percentage of schools with <90% vaccinated in 2017-18
Only schools that reported data
```{r}
  mmr%>% filter(yr==2018, mmr_pocket!='not reported') %>% tabyl(grade, mmr_pocket)%>%
   adorn_percentages("row") %>%
  adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  adorn_ns() %>% 
  adorn_title("top")%>%
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

#Percent of schools with <90% vaccinated, 2012-13
```{r}
  mmr%>% filter(yr==2013, mmr_pocket!='not reported') %>% tabyl(grade, mmr_pocket)%>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  adorn_ns() %>% 
  adorn_title("top")%>%
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

#By type of school- Kindergarten
2017-18 school year
```{r}
  mmr%>% filter(yr==2018, mmr_pocket!='not reported',  grade=='Kindergarten') %>%  tabyl(type2, mmr_pocket)%>%

  adorn_percentages("row") %>%
   adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  adorn_ns() %>% 
  adorn_title("top")%>%
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

#By type of school- Seventh
2017-18 school year
```{r}
  mmr%>% filter(yr==2018, mmr_pocket!='not reported',  grade=='Seventh') %>%  tabyl(type2, mmr_pocket)%>%

  adorn_percentages("row") %>%
   adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  adorn_title("top")%>%
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

#Kindergarten, 2017-18, by reason
Main factor in why vax rate is low<br>
in progress means lots of kids partially vaccinated or incomplete paperwork<br>
non-medical waiver means lots of kids with non-medical waiver<br>
equal means both things are pushing down the vax rate, equally<br>

```{r}
mmr %>% filter(yr==2018, mmr_pocket=='yes', grade=='Kindergarten') %>% 
  group_by(pocket_reason) %>% 
  summarize(count=n()) %>% 
  mutate(pct=count/sum(count))
```



#Schools with <90% vaccinated, by reason and type
Seventh grade, 2017-18
```{r}
  mmr%>% filter(yr==2018, mmr_pocket=='yes', grade=='Seventh') %>%  tabyl( type2, pocket_reason)%>%
   adorn_totals("row") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  adorn_ns() %>% 
  adorn_title("top")%>%
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

#Schools with <90% vaccinated, by reason and type
Kindergarten grade, 2017-18
```{r}
  mmr%>% filter(yr==2018, mmr_pocket=='yes', grade==
                  'Kindergarten') %>%  tabyl( type2, pocket_reason)%>%
   adorn_totals("row") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  adorn_ns() %>% 
  adorn_title("top")%>%
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```



#Revised vaccination rates by school type
Both grades combined
```{r}
mmr %>% filter(yr==2018, mmr_pocket!='not reported') %>% 
  group_by(type2) %>% 
  summarise(complete = sum(complete_mmr), 
            partial = sum(inprogress_mmr),
            nonmed = sum(co_mmr),
            med = sum(me_mmr)) %>% 
  mutate(total = complete+partial+nonmed+med,
         pctcomplete=complete/total,
         pctwaiver = nonmed/total) %>% 
    knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")

```

#Vaccination and waiver rates by grade
2017-18

```{r}
mmr %>% filter(yr==2018, mmr_pocket!='not reported') %>% 
  group_by(grade) %>% 
  summarise(complete = sum(complete_mmr), 
            partial = sum(inprogress_mmr),
            nonmed = sum(co_mmr),
            med = sum(me_mmr)) %>% 
  mutate(total = complete+partial+nonmed+med,
         pctcomplete=(complete/total)*100,
         pctwaiver = (nonmed/total)*100) %>% 
  select(grade, pctcomplete, pctwaiver) %>% 
    knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```


#Vaccination and waiver rates by school type
Kindergarten 2017-18

```{r}
mmr %>% filter(yr==2018, mmr_pocket!='not reported', grade=='Kindergarten') %>% 
  group_by(type2) %>% 
  summarise(complete = sum(complete_mmr), 
            partial = sum(inprogress_mmr),
            nonmed = sum(co_mmr),
            med = sum(me_mmr)) %>% 
  mutate(total = complete+partial+nonmed+med,
         pct=complete/total,
         pctwaiver=nonmed/total) %>% 
    knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

#Vaccination and waiver rates by type
Seventh grade, 2017-18
```{r}
mmr %>% filter(yr==2018, mmr_pocket!='not reported', grade=='Seventh') %>% 
  group_by(type2) %>% 
  summarise(complete = sum(complete_mmr), 
            partial = sum(inprogress_mmr),
            nonmed = sum(co_mmr),
            med = sum(me_mmr)) %>% 
  mutate(total = complete+partial+nonmed+med,
         pct=complete/total,
         pctwaiver=nonmed/total) %>% 
    knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```




#Percent of schools by poverty/income - 2018
1=low poverty (below 25%)<br>
4=high poverty (above 75%)<br>
mmr_pocket means school has less than 90% vaccinated<br>
Includes both kindergarten and 7th grade<br>
Only public schools
```{r}
  mmr%>% filter(yr==2018, income_bucket!='NA') %>%  tabyl(mmr_pocket, income_bucket)%>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  adorn_title("top")%>%
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

#Diversity 
Only includes public schools<br>
Kindergarten and 7th grade
```{r}
  mmr%>% filter(yr==2018, diversity_bucket!='NA') %>%  tabyl(mmr_pocket, diversity_bucket)%>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  adorn_title("top")%>%
  knitr::kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F, position="left")
```

#export data for graphics
```{r}

graphicsdata <-  mmr %>% filter(mmr_pocket!='not reported') %>%  group_by(grade, yr, type2, mmr_pocket) %>%     summarise(count=n(), complete = sum(complete_mmr), 
            partial = sum(inprogress_mmr),
            nonmed = sum(co_mmr),
            med = sum(me_mmr))

write.csv(graphicsdata, 'graphicsdata.csv', row.names=FALSE)
```



```{r}
childcare_vax_foronline %>% filter(MMR_pct_vac<.95, enrollment_2yrsup>0) %>% summarise(count=n())
```

```{r}
mmr %>% filter(yr==2018, grade=='Kindergarten') %>% 
group_by(mmr_pocket) %>% 
  summarise(enr= sum(enroll_new), count=n()) %>% 
  mutate(pct=enr/sum(enr), pct2=count/sum(count))
```

